---
title: "Prepare stomach content data for Lammska analysis"
author: "Max Lindmark & Viktor Thunell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    embed-resources: true
    fig-width: 8
    #fig-asp: 0.618
knitr: 
  opts_chunk:
    fig.align: center
editor: source
execute: 
  echo: true
  eval: true
  cache: false
---

## Load libraries

```{r libs}
#| message: false
#| warning: false
library(tidyverse)
library(tidylog)
library(janitor)
library(devtools)
library(sdmTMB)
library(patchwork)
library(terra)
library(mapplots)

# Import some plotting functions
# Source code for map plots
# You need: devtools::install_github("seananderson/ggsidekick") # not on CRAN; library(ggsidekick)
devtools::source_url("https://raw.githubusercontent.com/VThunell/Lammska_cod-fr/main/R/functions/map-plot.R")
options(ggplot2.continuous.colour = "viridis")

# Set path
home <- here::here()
```

## Read data

I downloaded all stomach data from the Baltic, all countries (year 1963-2021), at 16:21  Jan 15 2024. This script prepares those, and two additional data sets, and combines them and prepares them for anlysis.

I want to analyse the individual level predator (cod) prey feeding ration, i.e weight total stomach content and specific prey types (e.g. saduria, herring and sprat) to cod. Hence we need, per cod stomach, the total prey weight and the sopecific weights of these species, and if they are absent, weight should be 0 . We also need coordinates for each cod. To get there, we need to do the following:

1. Read in data
    a. downloaded data from ICES stomach database (new data base)
    b. complete new with old database (ODB) data (since large parts have still not been uploaded)
    c. add additional newer (SE) data and make data set with one row per prey item
2. Some renaming and data cleaning
3. Filter predators with the above prey present and calculate the total weight of these for each predator (one row per predator)
4. Filter predators where these prey are not present. VT??
5. Bind rows, and replace NA with 0 VTY??


```{r read data}
# New database data
fi <- read_csv(paste0(home, "/data/stomach/StomachContent_0115213707/File_information.csv"))
hi <- read_csv(paste0(home, "/data/stomach/StomachContent_0115213707/HaulInformation.csv"), 
               col_types = list(
                 tblUploadID = col_double(),
                 tblHaulID = col_double(),
                 Ship = col_character(),
                 Gear = col_character(),
                 HaulNo = col_double(),
                 StationNumber = col_double(),
                 Year = col_double(),
                 Month = col_double(),
                 Day = col_double(),
                 Time = col_character(),
                 ShootLat = col_double(),
                 ShootLong = col_double(),
                 HaulLat = col_double(),
                 HaulLong = col_double(),
                 ICESrectangle = col_character(),
                 Depth = col_double(),
                 Survey = col_character(),
                 ICESDatabase = col_character(),
                 Notes =  col_character())) # the parsing issues of hi causes no problems but specifying col_types corrects the datatypes. 
pred <- read_csv(paste0(home, "/data/stomach/StomachContent_0115213707/PredatorInformation_edit.csv")) # edited in text editor (not excel or equivalent) to remove three instances of erroneous '"' in column 'Notes' at row 7139,7144 and 7296.
prey <- read_csv(paste0(home, "/data/stomach/StomachContent_0115213707/PreyInformation.csv"))

# Old database data (ODB)
shimmy <- read_csv(paste0(home, "/data/stomach/StomachDataOld.csv")) 

# New SE data
#SE <- read_csv(paste0(home, "/data/stomach/aggregated_stomach_data.csv")) 
SE <- read_csv(paste0(home, "/data/stomach/full_stomach_data.csv"))

# BITS survey data to improve coordinates in ODB
# bits_hh <- getDATRAS(record = "HH", survey = "BITS", years = 1963:2022, quarters = c(1,2,3,4))
# write_csv(bits_hh, paste0(home, "/data/DATRAS_exchange/bits_hh.csv"))
bits_hh <- read_csv(paste0(home, "/data/DATRAS_exchange/bits_hh.csv"))

```
## New database data (dowloaded Jan 15 2024)
Start with new database data. The description of the data can be found here: <http://datsu.ices.dk/web/selRep.aspx?Dataset=157>

```{r glimpse new data}
names(fi)
names(hi)
names(pred)
names(prey)
```

## Join all data files (new db)
We do this in a specific order: fi -> hi -> pred -> prey.

For some joins, there are multiple column names shared in addition to the key that I could remove and keep only the ID key and the non-shared columns but instead I keep them. First I need to ensure they are the same (VT what is the same and how do you do this?), and not only have the same name. Will also check if both datasets have the same amount of NA before choosing which column to carry from which dataset.

```{r}
hifi <- left_join(hi, fi, by = "tblUploadID") # join hail data with file info

comcol_hifi_pred <- intersect(colnames(pred), colnames(hifi))

# Check if any of the two datasets have NA in the common columns
unique(is.na(hifi |> dplyr::select(all_of(comcol_hifi_pred))))
unique(is.na(pred |> dplyr::select(all_of(comcol_hifi_pred))))

# The column Notes does have NAs and different meanings in hi and fi so we will remove Notes before joining
pred <- left_join(pred |> dplyr::select(-Notes), # join in predator data
                  hifi |> dplyr::select(-Notes),
                  by = comcol_hifi_pred[!comcol_hifi_pred == "Notes"])
```

Now let's join predator data to prey data following the same procedure.

```{r}
intersect(colnames(pred), colnames(prey))

# Length is a common column, but it corresponds to predator or prey. Rename!
pred <- pred |> rename(pred_length = Length)
prey <- prey |> rename(prey_length = Length)

comcol_prey_pred <- intersect(colnames(pred), colnames(prey))

# Check if any of the two datasets have NA in the common columns
unique(is.na(pred |> dplyr::select(all_of(comcol_prey_pred))))
unique(is.na(prey |> dplyr::select(all_of(comcol_prey_pred))))

# Rename "Notes" in prey data to avoid confusion as to which dataset it belongs to
prey <- prey |> rename(prey_notes = Notes)

# In the Swedish data, 2 means regurgitated and 1 is intact, but for the rest, 1 means regurgitated, 0 or NA means intact. All NA values are from 2020 and 2021 and for all regurguitated=0 before 2005 (see Neuenfeldt 2020 discussion) we dont know if they are truly not regurgitated. Therefore there is not much info of value from the regurgitated column. 
pred <- pred |>
  mutate(Regurgitated_st = Regurgitated,
         Regurgitated_st = ifelse(Country == "SE" & Regurgitated_st == 2, 1, Regurgitated_st))

# Regurgitated stomachs are only about 2% and come from 2021, 2018 and 2013 (and 1 in 1981). The NAs are from 2020 and 2021.
pred |>
  group_by(Year, Regurgitated_st) |>
  summarize(count = n())

# There are regurgitated stomachs (column in pred) have information in prey info, i.e. either its incorrect or they have signs of regurgitation but prey in the stomach. These are various prey types from 2018 to 2021. Lets remove these for the fr analyses
regurg_ids <- pred |>
  filter(Regurgitated_st == 1) 
prey |> filter(tblPredatorInformationID %in% regurg_ids$tblPredatorInformationID) |> distinct(Weight, AphiaIDPrey, Year) |> as.data.frame()
 
#Remove regurgitated stomachs from pred and prey to reduce issues with the predator prey ratio becoming incorrect.
pred <- pred |>
  filter(!tblPredatorInformationID %in% regurg_ids$tblPredatorInformationID)
prey <- prey |>
  filter(!tblPredatorInformationID %in% regurg_ids$tblPredatorInformationID)

# Join in pred info into prey. There are rows only in pred (y) that are not in prey. These are either empty, regurgitated or incorrect. Empties will be added later.
d <- left_join(prey, pred, by = c(comcol_prey_pred))

```

## Calculate total weight of specific prey species by unique predator ID

For NA and zero weights, we estimate weight if length is present. If length is not `NA` and Weight is 0 or NA, estimate weight based on length and Count. Else give weight `NA` and drop it. Because these are not true empty, else there wouldn't be species-information

```{r}
# Estimate weight based on count. In the data, if count is >1, the weight is grouped.  In some cases, all of Weight, Count and prey_length is NA, i.e. there are presences of prey but no information for calculating the total weight or there is a length but no count (0) which we will treat as one and retrieve a weight, I think this is less wrong.

# NA length unit matches the number of missing lengths. Where we don't have a unit is where length is lacking.
sum(is.na(d$UnitLngt))/length(d$UnitLngt)
sum(is.na(d$prey_length))/length(d$prey_length)

# Calculate weight of prey given count and prey length
d_c <- d |> 
  mutate(prey_length = if_else(UnitLngt == "mm" & prey_length >= 0, prey_length/10, prey_length, missing = NA),
         Weight = replace_na(Weight, -9),
         Count = replace_na(Count, -9),
         prey_length = replace_na(prey_length, -9), # Because there is both negative values and NAs
         Count = if_else(prey_length > 0 & Count <= 0, 1, Count, missing = -9),# When prey_length exists but Count <= 0 we set them to 1.
         prey_weight_source = if_else(Weight <= 0 & prey_length > 0 & Count >= 0 , "estimated", "observed", missing = NA),
         Weight = if_else(prey_weight_source == "estimated", (0.01*prey_length^3)*Count, Weight, missing = -9),
         ind_weight = if_else(Weight > 0 & Count > 0, Weight / Count, NA)) # Calcuate average prey weights

d_c |> # some unreasonable lengths but not too much
  ggplot(aes(Weight)) +
  geom_histogram() +
  facet_wrap(~AphiaIDPrey, scales = "free")
  
d_c |> filter(Weight <= 0) |> dplyr::select(AphiaIDPrey, Weight, prey_length, Count)
# Even if we estimate the prey weight based on the length of the prey and the number of the prey, we still have 173 rows that Weight == -9 or 0. Some of these are due to missing AphiaID, e.g. plastic and stone. We can give them the average weight. This may produce outliers when the predator is small compared to the avg prey size and but these would be removed in a later cleaning stage. Lets do this.

d_c |>
  filter(is.na(AphiaIDPrey))

d_c |> # Of course there are extreme weights that should not be included
  ggplot(aes(ind_weight)) +
  facet_wrap(~AphiaIDPrey, ncol = 3, scales = "free") +
  geom_histogram()

d_c |> # 893 g sprat? Doubt it! 
  filter(AphiaIDPrey %in% c(293743, 1625944, 126425, 322683, 236448),
         ind_weight > 200) |> as.data.frame() 

d_c |> # Saduria above 10 g, likely Count is wrong etc.
  filter(AphiaIDPrey %in% c(293511, 119034),  
         ind_weight > 10) |> as.data.frame() 
#VT: Many saduria is way off, mean weights of 1 g. max weight of 7,8 or 9? The count may be off etc. If the saduria analyses becomes important I need to revisit these values (https://doi.org/10.1111/j.0021-8790.2004.00800.x and https://www.jstor.org/stable/24831823 ). These outliers likely dissapear when I filter on feeding ratio. Can be calculated from Sharkweb data.

# To control for outliers affecting average of individual weight, remove outliers outside quartiles
prey_avg_ind_weight <- d_c |>
  filter(!is.na(AphiaIDPrey)) |>
  group_by(AphiaIDPrey) |> 
    filter(!ind_weight %in% boxplot.stats(ind_weight)$out) |>
    reframe(avg_weight = mean(ind_weight, na.rm = TRUE))

# This should be the average prey weight which we can use to calculate the weight of these prey if we have the counts. Left join that summarized data and do the estimate of weight based on length. But first figure out which unit prey size is

# Join average weight and estimate weight
d_c <- d_c |> 
  left_join(prey_avg_ind_weight) |> 
  mutate(Weight = if_else(Weight <= 0 & Count > 0 & !is.na(AphiaIDPrey), Count * avg_weight, Weight, missing = -9),
         prey_weight = replace_na(Weight, 0))

# Max it this make more sense to have zeros could be rounding errors because they are present, whereas NAs could be because it simply wasn't weighed). 
# Remove the negative weights (130 rows, <1%) and the NAs (6 rows, which are caused by NAs or lacking weight for an unusual prey)! Keep the ID's though because we also want to remove these from the full data set so that we don't inflate the total number of empty stomachs, because these are not actually empty. This removes 2 rows
ids_to_remove <- d_c |> filter(Weight <= 0 | is.na(Weight))
ids_to_remove |> as.data.frame() |> distinct(Weight, prey_length, Count)

d_c <- d_c |> 
  filter(!(Weight <= 0 & Count <= 0 & prey_length <= 0),
         !is.na(Weight))

```
Next we need to summarize our prey weights by predator and prey group. 

```{r}

length(which(is.na(d_c$AphiaIDPrey)))/length(d_c$AphiaIDPrey) # 4% NA, they will go into the other category

agg_prey <- d_c |> # isopoda are likely often saduria but may also be other species. 
  mutate(prey_class = NA,
         prey_class = if_else(AphiaIDPrey %in% c(120113,130735,101744,101156,104152,101537,107552,110487, 101160,103079,103077,119822,120136,103251,102101,120120,141579,140480,120115,101361,144129,104165,883,103068,104878,1078,104517,104241,1066,51,1135,234024,326114,149668,882,106782,106674,105,1248,101063,107614,152302,140430, 107032,138999,138081,102292,107613), "other_invert", prey_class),
         prey_class = if_else(AphiaIDPrey %in% c(293511, 119034), "saduria", prey_class), 
         prey_class = if_else(AphiaIDPrey %in% c(126928,127196,126436,126450,125999,254529,127141,127123,126916, 127204,126996,127203,125537,152352), "benth_fish", prey_class),
         prey_class = if_else(AphiaIDPrey %in% c(293743, 1625944, 126425, 322683, 236448), "sprat", prey_class),
         prey_class = if_else(AphiaIDPrey %in% 126417, "herring", prey_class),
         prey_class = if_else(AphiaIDPrey %in% c(125464, 125716), "unid_clupeids", prey_class),
         prey_class = if_else(AphiaIDPrey %in% c(126752, 126505, 126736, 125516), "other_fish", prey_class),
         prey_class = if_else(!prey_class %in% c("other_invert", "saduria", "benth_fish", "herring", "sprat", "other_fish", "unid_clupeids"), "other_or_NA", prey_class)) # NOTE "other_or_NA"

# same number NA in agg_prey as in d!
# fixing "unid_clupeids" below 
table(agg_prey$prey_class)

# Negative numbers are likely because they couldn't be counted only weighed
min(prey$Count, na.rm = TRUE)
min(prey$Weight, na.rm = TRUE)

agg_prey |> filter(Weight == 0)
```

With these estimates of weight based on either length or worst case, average weight of that prey, we calculate the total weight of these prey per individual predator stomach, and then pivot wider. 

```{r}

agg_prey_summed <- agg_prey |> 
  group_by(tblPredatorInformationID, prey_class) |> 
  summarise(tot_weight = sum(Weight)) |> 
  ungroup()

sum(is.na(agg_prey_summed$prey_class))

agg_prey_summed <- agg_prey_summed |> 
  pivot_wider(names_from = "prey_class", values_from = "tot_weight", values_fill = 0) 

# Split unidentified clupeids on sprat and herring  
agg_prey_summed <- agg_prey_summed |> 
  mutate(sprat = sprat + unid_clupeids*0.5,
         herring = herring + unid_clupeids*0.5) |> 
  dplyr::select(!unid_clupeids)

```
Next I will `left_join` in the remaining predator information, and after that `bind_rows` "empty stomachs" (with respect to these 3 prey species). Since the IDs are not overlapping, it doesn't matter that I already have some 0's here for some species

```{r}
agg_prey_summed <- 
  agg_prey_summed |> 
  left_join(pred, by = "tblPredatorInformationID") |> 
  drop_na(pred_length) 

t <- agg_prey_summed |>
  dplyr::select("other_invert", "benth_fish", "other_fish", "sprat", "herring", "other_or_NA") |>
  rowSums() == 0
length(which(t)) # empty stomachs..
```
Now add in the "empty stomachs" from pred using bind_rows. When I `bind_rows`, the columns that are not matching get `NA`. The only column not matching should be the average weight columns. They will get `NA`, and I'll change it to 0.

```{r}
# The empty stomachs to be added are those that are empty in predator data. For data before 2005 we don't know whether they are regurgitated or empty because of not eating. For an analysis using all data (1963-) we need to keep that in mind.
# Because of the issue of not getting prey weight even though they are present above, we need to make sure to drop these stomachs in the full data set also before joining, so that we don't treat them as empty!

# StomachEmpty = 1 is an empty stomach even though this is not an mandatory paramter (?). Since regurgitated==1 and NA have been removed, all empty stomachs seem to be truly empty, i.e. there are no tblpredatorinformationIDs with stomach empty that are in the prey data (No overlap in stomachs between the predator and prey data sets).
empty_stom <- pred |> 
 filter(StomachEmpty == 1) |>
 filter(!tblPredatorInformationID %in% unique(agg_prey_summed$tblPredatorInformationID))

# Bind rows!
dd <- agg_prey_summed |> 
  bind_rows(empty_stom)

# Denmark report in kg
ggplot(dd, aes(IndWgt)) + 
  facet_wrap(~Country, scales = "free") + 
  geom_histogram()

dd <- dd |> 
  mutate(IndWgt = ifelse(Country %in% c("DK"), IndWgt*1000, IndWgt))

dd |> # Note the one data point in 1989
  ggplot(aes(IndWgt)) + 
  facet_wrap(~Year, scales = "free") +
  geom_histogram()

# Sweden report in mm
dd |> # Note the one data point in 1989
  ggplot(aes(pred_length)) + 
  facet_wrap(~Country, scales = "free") +
  geom_histogram()

dd <- dd |> 
  mutate(pred_length = ifelse(Country %in% c("SE"), pred_length/10, pred_length))

# Yes, works as intended, see the added NAs in the selected prey weights. Make them 0!
dd <- dd |> 
  mutate(other_or_NA = replace_na(other_or_NA, 0),
         sprat = replace_na(sprat, 0),
         herring = replace_na(herring, 0),
         saduria = replace_na(saduria, 0),
         other_invert = replace_na(other_invert, 0),
         benth_fish = replace_na(benth_fish, 0),
         other_fish = replace_na(other_fish, 0))

# Remove non cod predators (e.g. Whiting 126438)
dd <- dd |> filter(AphiaIDPredator %in% c(126436))

ggplot(dd, aes(pred_length, IndWgt)) + # 0, negative or NAs weight with lengths
  geom_point()

# Estimate predator weight if we have their length?
dd <- dd |> 
  mutate(IndWgt = replace_na(IndWgt, 0),
         pred_weight_source = ifelse(IndWgt <= 0, "estimated", "observed"),
         IndWgt = ifelse(pred_weight_source == "estimated", 0.01*pred_length^3, IndWgt))

dd |> 
  arrange(desc(pred_weight_source)) |>
  ggplot(aes(pred_length, IndWgt, col = pred_weight_source)) +
  geom_point(shape = 4)

```

## Old database data

The data in the old ICES database contains much more data than what is found in the new version of the database but is in poor shape containing many errors and unique identiifers are missing. Below, identifiers for hauls and predators are made and the old db is cleaned.

```{r odb general}
glimpse(shimmy)
# Each row in old data contains a prey item, i.e. we have many rows for each predator and there is no unique predator ID. We need to generate an identifier so that we can compare observations the old and new data. The ICESsampleID is for the predator I assume, ICESitemID are prey items but there are duplicates, possibly mainly due to the two data sets (Year of stomach and Stomach tender d using the same IDs which become duplicates?).  
# By creating an ID with Date (Year, Month and Day), Country and Haul (HaulNo (new data) and Haul (old data)) we can identify the data missing from the new database. Ship info in the old data is unclean, by instead assuming that Country should be unique to date and haul (sort of Country as a proxy for Ship). However, there are 41506 rows where country and ship is missing. VT removes these. 
#length(is.na(old_db[which(is.na(old_db$Country)),]$Ship))

old_db <- janitor::remove_empty(shimmy, which = "cols") # remove cols with only Nas to increase readability

# fix names
old_db <- old_db |>
  mutate(Country = if_else(Country == "Latvia", "LV",
                            if_else(Country == "Denmark", "DK", 
                                   if_else(Country == "Poland", "PL", 
                                          if_else(Country == "Sweden", "SE",
                                                 if_else(Country == "Germany", "DE", "NotBaltic")))),
                           missing = "NotBaltic"))

length(which(is.na(old_db$Country)))
length(which(is.na(old_db$Haul)))
length(which(is.na(old_db$ICES_StatRec)))

# we get rid of almost all NAs (198) in Haul by excluding based on lat lon (Country not needed then)
# filter and create identifier
old_data <- old_db |> 
  rename(HaulNo = Haul,
         ICESrectangle = ICES_StatRec) |> 
  filter(Predator_AphiaID %in% c(126436), # remove non cods
         between(Latitude, 53, 65.5) & between(Longitude, 10, 24)) |> # remove data outside the Baltic
  mutate(Haul_ID = paste(Country, Year, Month, Day, HaulNo, ICESrectangle, sep = "_"))

# join hi and fi from new db to get columns for identifier
# add Country and ICESrectangle to new db prey data from haul data and create identifier
prey_temp <- prey |>
  left_join(dplyr::select(hifi, tblUploadID, Country, ICESrectangle), by = "tblUploadID", multiple = "any") |>
  mutate(Haul_ID = paste(Country, Year, Month, Day, HaulNo, ICESrectangle, sep = "_"))

# identify cod in hauls in the baltic not present in the new data. I.e. what the old db is adding to the new db
missing_db <- old_data |>
  anti_join(prey_temp, by = "Haul_ID") # return all rows from x (old_d) without a match in y (prey)
```


```{r odb prey pred lenghts ans weights}
# fix predator lengths and weights. Esitmate weight when length exists
missing_db2 <- missing_db |> 
  rename(pred_length = `Predator_Lengh(mean)`,
         pred_weight = `Predator_Weight(mean)`) |>
  mutate(pred_weight_source = if_else(is.na(pred_weight) | pred_weight <= 0, "estimated", "observed"),
         pred_weight = if_else(pred_weight_source == "estimated", (0.01*pred_length^3), pred_weight, missing = NA)) 

missing_db2 |> # 819 prey weights out of 71 000 should be possible to fix
  select(Prey_LowerLengthBound, Prey_UpperLengthBound, Prey_Weight, Prey_TotalNo) |>
  filter(is.na(Prey_Weight) | Prey_Weight == 0) |>
  filter((Prey_LowerLengthBound > 0 | Prey_UpperLengthBound > 0))

# fix prey lengths and weights. Estimate weight from length and count. If count is 0 or NA when length > 0, assume count=1.
missing_db2 |>  # 70275 zeros out of which many are NA
  filter(Prey_Weight > 0) |>
  summarise(n()) 

missing_db2 <- missing_db2 |> 
  rename(count_prey = Prey_TotalNo,
         prey_weight = Prey_Weight) |>
  mutate(Prey_UpperLengthBound = replace_na(Prey_UpperLengthBound, 0),
         Prey_LowerLengthBound = replace_na(Prey_LowerLengthBound, 0),
         prey_length = if_else(Prey_UpperLengthBound == 0, Prey_LowerLengthBound, 
                               (Prey_UpperLengthBound + Prey_LowerLengthBound)/2, missing = NA),# there were few Upper (233)
         count_prey = replace_na(count_prey, 0),
         prey_weight = replace_na(prey_weight, 0),
         prey_weight_source = if_else(prey_weight > 0, "observed",
                                      if_else(count_prey > 0 & prey_length > 0, "estimated",
                                              if_else(count_prey == 0 & prey_length > 0, "estimated_zerocount", 
                                                      NA, missing = NA), 
                                              missing = NA)),
         prey_weight = if_else(prey_weight_source == "estimated_zerocount", (0.01*prey_length^3)*1, 
                               if_else(prey_weight_source == "estimated", (0.01*prey_length^3)*count_prey, 
                                       prey_weight, missing = 0), 
                               missing = 0), 
         prey_weight_ind = if_else(count_prey > 0 & prey_weight > 0, prey_weight / count_prey, NA))

table(missing_db2$prey_weight_source, useNA = "ifany")
missing_db2 |>  # 70275 zeros out of which many come from NAs
  filter(prey_weight <= 0)

# What to do with 70 000 missing weights. We can give them average weight like for the new db. 
missing_db2 |> # where there is a prey species IDd, there are only 611 zero weights that we can fix. Here, I assume count Na/0 is zero. 
  filter(!is.na(Prey_AphiaID) & prey_weight == 0 & count_prey >= 0)

length(unique(prey_avg_ind_weight_odb$Prey_AphiaID))

# To control for outliers affecting average of individual weight, remove outliers outside quartiles
prey_avg_ind_weight_odb <- missing_db2 |> #prey_avg_ind_weight_odb |>
  filter(!is.na(Prey_AphiaID)) |> 
  group_by(Prey_AphiaID) |> 
    filter(!prey_weight_ind %in% boxplot.stats(prey_weight_ind)$out) |>
    reframe(avg_weight = mean(prey_weight_ind, na.rm = TRUE))

# Join average weight and estimate weight
missing_db3 <- missing_db2 |> 
  left_join(prey_avg_ind_weight_odb) |> 
  mutate(prey_weight = if_else(prey_weight <= 0 & count_prey > 0 & !is.na(Prey_AphiaID), count_prey * avg_weight, prey_weight))

missing_db3 |>  # 70275 zeros out of which many come from NAs
  #filter(prey_weight == 0)
  filter(is.na(prey_weight))

# remove zero weight predators and Id those with empty stomachs, make predator identifier. Rempove NA prey weights
missing_db3 <- missing_db3 |> 
  filter( pred_weight > 0,
          !is.na(prey_weight)) |>
  mutate(pred_ID = paste(Haul_ID, `SampleNo(FishID)`, sep = "_"),
         empty_stom = ifelse(prey_weight <= 0 | is.na(prey_weight), 1, 0))


```


```{r prey classes}

# Make prey classes
missing_db4 <- missing_db3 |>
  mutate(prey_class = NA, 
         prey_class = if_else( Prey_LatinName %in% c("Mysis mixta", "Bylgides sarsi", "Crangon crangon", "Pontoporeia affinis", "Pontoporeia femorata", "Gammarus", "Mysidae",  "Halicryptus spinulosus", "Neomysis integer", "Hyperia galba",  "Gastropoda",  "Priapulida", "Priapulidae", "Priapulus caudatus", "Caprellidae",  "Gammaridae", "Hyperiidae", "Corophium volutator", "Bathyporeia pilosa", "Monoporeia affinis", "Macoma balthica", "Diastylis rathkei", "Bivalvia", "Crustacea", "Brachyura", "Caridea", "Paguridae", "Leucosiidae", "Hippolytidae", "Crangonidae", "Pandalidae", "Pasiphaeidae", "Processidae", "Galathea intermedia", "Nephrops norvegicus",  "Liocarcinus depurator", "Palaemon elegans", "Pandalus montagui", "Pasiphaea multidentata", "Pasiphaea sivado", "Ostracoda", "Copepoda", "Meganyctiphanes norvegica", "Decapoda", "Isopoda", "Amphipoda", "Cumacea", "Idotea balthica", "Gastrosaccus spinifer", "Mysis oculata", "Mysis relicta", "Praunus flexuosus", "Praunus inermis", "Ophiuridae", "Bylgides",  "Aphrodita aculeata",  "Scoloplos armiger", "Insecta", "Mytilus edulis", "Synchaeta", "Scyphozoa", "Aurelia aurita", "Hydrobia", "Loligo vulgaris", "Mya arenaria", "Hediste diversicolor", "Cardium edule", "Terebellides stroemi", "Nereididae",  "Pestarella tyrrhena", "Mollusca", "Polybiidae", "Annelida", "Polychaeta", "Amphibalanus improvisus"), "other_invert", prey_class),
         prey_class = if_else( Prey_LatinName %in% c("Saduria entomon"), "saduria", prey_class),
         prey_class = if_else( Prey_LatinName %in% c("Enchelyopus cimbrius", "Gadus morhua", "Zoarces viviparus", "Pomatoschistus minutus", "Platichthys flesus", "Pomatoschistus", "Myxine glutinosa", "Lampetra fluviatilis", "Pleuronectiformes", "Ammodytes tobianus", "Pisces",  "Gadidae", "Ammodytidae", "Callionymidae", "Gobiidae", "Stichaeidae", "Zoarcidae", "Pleuronectidae", "Cottidae", "Melanogrammus aeglefinus", "Merlangius merlangus", "Pollachius virens", "Spinachia spinachia", "Hyperoplus lanceolatus", "Aphia minuta", "Neogobius melanostomus", "Pomatoschistus microps", "Pholis gunnellus", "Trachinus draco", "Hippoglossoides platessoides", "Limanda limanda", "Pleuronectes platessa", "Cottus gobio", "Myoxocephalus scorpius", "Taurulus bubalis", "Cyclopterus lumpus", "Maurolicus muelleri", "Sander lucioperca", "Rutilus rutilus",  "Lumpenus lampretaeformis", "Myoxocephalus quadricornis", "Vimba vimba"), "benth_fish", prey_class),
         prey_class = if_else( Prey_LatinName %in% c("Sprattus sprattus"), "sprat", prey_class),
         prey_class = if_else( Prey_LatinName %in% c("Clupea harengus"), "herring", prey_class),
         prey_class = if_else( Prey_LatinName %in% c("Clupea", "Clupeidae"), "unid_clupeids", prey_class),
         prey_class = if_else( Prey_LatinName %in% c("Gasterosteus aculeatus", "Osmerus eperlanus", "Belone belone", "Engraulis encrasicolus"), "other_fish", prey_class),
         prey_class = if_else(!prey_class %in% c("other_invert", "saduria", "benth_fish", "herring", "unid_clupeids", "sprat", "other_fish"), "other_or_NA", prey_class, missing = NA)) # NOTE, "clupea" and"clupeidea" etc. are "other"

# Duplicates?
dup <- missing_db4 |>
  group_by(Haul_ID) |> 
  dplyr::select(-Dataset, -ICES_SampleID) |> #ID, `Predator_Weight(mean)`) |> 
  duplicated() |> 
  which()

# remove duplicates
missing_db4 <- missing_db4[-dup,]

# in the old db,, we have both pred and prey data. This means we need to summarize and then add in pred info from the whole db df. To keep the columns we need we use mutate instead of summarise(). 
length(unique(missing_db4$pred_ID)) 

missing_db4_summed <- missing_db4 |> 
  #filter(empty_stom == 0) |> 
  group_by(pred_ID, prey_class) |> 
  summarise(tot_weight = sum(prey_weight)) |> 
  #mutate(tot_weight = sum(prey_weight)) |>
  ungroup()

missing_db4_summed <- missing_db4_summed |> 
  pivot_wider( names_from = "prey_class", values_from = "tot_weight", values_fill = 0) # `values_fill = 0` make 0 not NA

# Split unidentified clupeids on sprat and herring  
missing_db4_summed <- missing_db4_summed |> 
  mutate(sprat = sprat + unid_clupeids*0.5,
         herring = herring + unid_clupeids*0.5) |> 
  dplyr::select(!unid_clupeids)

# colnames(missing_db3) selecting columns without prey info in the join.
mm <- missing_db4_summed |> 
   left_join( dplyr::select(missing_db3, Dataset, Country, Ship, Latitude, Longitude, Estimated_Lat_Lon, ICESrectangle, Year, Month, Day, Time, Station, HaulNo, Depth, 'SampleNo(FishID)', Predator_AphiaID, Predator_LatinName, pred_weight, pred_length, prey_weight, Haul_ID, pred_weight_source, prey_weight_source, pred_ID, empty_stom), by = "pred_ID", multiple = "first")

mm |> # there are unlikeley extremes. These are filtered on feeding ratio in the analysis.   
  rowwise() |>
  mutate(total = sum(other_or_NA, other_invert, saduria, benth_fish, herring, sprat, other_fish)) |>
  filter( total > 1000) |>
  arrange(desc(total))

mm |> count(empty_stom) # 36 % empty stomachs

mm |> 
  arrange(desc(pred_weight_source)) |>
  ggplot(aes(pred_length, pred_weight, col = pred_weight_source)) +
  geom_point(shape = 4)
```


```{r join in coordinates from BITS}
# # one third of the data has estimated the coordinates  according to Estimated_Lat_Lon
sum(old_db$Estimated_Lat_Lon == "Yes", na.rm = TRUE)/nrow(old_db) 

mm |> # But its likely all old db data since the center of the ices rectangle equals the lat lon data
     # mutate(Latitude = round(Latitude, 2),
     #        Latitude = round(Latitude, 2)) |> 
     filter(ices.rect(ICESrectangle)$lon == Longitude & ices.rect(ICESrectangle)$lat == Latitude) |>
     count()

length(unique(mm$Longitude)) # only 12 unique lats 
mm |> # and 70 unique coords in the old db
  select(Latitude, Longitude) |> 
  distinct()

# Make an identifier
intersect(tolower(colnames(mm)), tolower(colnames(bits_hh)))
length(which(is.na(mm$HaulNo))) # a few rows lacks haul number.

mm <- mm |> 
  mutate(Hid = paste(Country, Year, Month, Day, HaulNo, ICESrectangle, sep = "_")) #

bits_hh <- bits_hh |> 
  mutate(Hid = paste(Country, Year, Month, Day, HaulNo, StatRec, sep = "_"))

length(intersect(mm$Hid, bits_hh$Hid) ) # 553 unique hauls that are matching based on Country, Year, Month, Day, ices_rect

mm |> # We should have done better in the current millenium... 
  filter(!Hid %in% unique(bits_hh$Hid)) |>
  mutate(decade = round(Year/10) * 10) |>
  group_by(decade, Country) |>
  count()

# Now I add lat/lon values from bits_hh to replace those in mm
mm <- mm |> 
  left_join(bits_hh |> dplyr::select(ShootLat, ShootLong, Hid)) |>
  mutate(lat = ifelse(is.na(ShootLat), Latitude, ShootLat),
         lon = ifelse(is.na(ShootLong), Longitude, ShootLong))

# now there is 591 unique coordinates
mm |>
  select(lat, lon) |> 
  distinct()
```

## New SE data

```{r}
# Calculate total weight of prey by predator ID and prey species (i.e., across prey sizes). First create wide data frame so that I can sum easily across prey groups (columns)
SE_wide <- SE %>% 
  group_by(pred_id, prey_latin_name) %>% 
  summarise(tot_prey_weight_g = sum(prey_weight_g)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = prey_latin_name, values_from = tot_prey_weight_g) %>% 
  mutate_all(~ifelse(is.na(.), 0, .)) %>% 
  clean_names()

# Now make some calculations and aggregate to some taxonomic level. Since all columns are assigned to some higher level group (or the same group), the sum of these is the total stomach content. Note that I have one group for unidentified clupeids, but also sprat and herring. So if I want the total of some aggregated group, then I need to add all the sub-groups.

SE_wide2 <- SE_wide %>% 
  mutate(other_invert = gammarus_sp + monoporeia_affinis + amphipoda + hyperia_galba + bivalvia + mytilus_sp + mytilus_sp_2 + mya_arenaria + macoma_balthica + mytilus_edulis + limecola_balthica + limecola_balthica_2 + mysidae + neomysis_integer + mysis_mixta + mysida + gastrosacus + pontoporeia_femorata + pontoporeia_femotara + crangon + crangon_crangon + idotea_balthica + cumacea + idotea_sp + praunus_flexuosus + crustacea + diastylis_rathkei + palaemon_sp + palaemon_elegans + caridea + copepoda + pontoporeiidae + decapoda + pontoporeidae + halicryptus_spinulosus + halicryptus_spinulosus_2 + priapulus + priapulus_caudatus + priapulus_caudatus_2 + hydrobia_sp + priapulida + halicryptus + halicryptus_spinolusus + priapulida_2 + prapulida + bylgides_sarsi + scoloplos_armiger + hediste_diversicolor + phyllodocida + polychaeta + pectinaria_sp + nephtys_ciliata,
         unid_clupeids = clupeidae + clupeidae_2,
         herring = clupea_harengus + clupea_harengus_2,
         saduria = saduria_entomon,
         sprat = sprattus_sprattus + sprattus_sprattus_2,
         benth_fish = gadus_morhua + gobiidae + gobiidae_2 + gobius_niger + neogobius_melanostomus + aphia_minuta + platichthys_flesus + enchelyopus_cimbrius + zoarces_viviparus + pleuronectidae + agonus_cataphractus + myoxocephalus_scorpius,
         other_or_NA = stone + stone_2 + plastic + plastics + sand + wood + carbon + carbon_2 + wood_2 + litter_plastics + sand_2 + algae + aglae + waste + remains + remains_2 + digestive_tract + mucus + mucus_2 + na,
         other_fish = pisces + pisces_2 +  + engraulis_encrasicolus + gasterosteus_aculeatus + scales + scales_2 + scomber_scombrus + pungitius_pungitius + ammodytidae + gasterosteidae)

SE_wide2 <- SE_wide2 |> # split clupeids into equal parts sprat and herring
  mutate(sprat = sprat + unid_clupeids*0.5,
         herring = herring + unid_clupeids*0.5) |> 
  dplyr::select(!unid_clupeids)

# Check it's correct! The sum of all groups should be the sum of all categories
t <- SE_wide2 |>
  rowwise() |> 
  mutate(all_cat_sum = sum(c_across(2:102)),
         group_sum = sum(c_across(103:109))) |>
  dplyr::select(all_cat_sum, group_sum) |>
  filter(!all_cat_sum == group_sum)  
which(!(t[1]- t[2]) < 10^-13) # only differences due to rounding.

# Select only aggregated columns 
SE_wide3 <- SE_wide2 %>%
  dplyr::select(pred_id, other_invert, herring, saduria, sprat, benth_fish, other_or_NA, other_fish)

# Add back in other information about the predator ID and add and rename some columns
ll <- SE |>
  dplyr::select(predator_latin_name, species, pred_weight_g, pred_length_cm,
                year, quarter, month, ices_rect, subdiv, haul_id, smhi_serial_no,
                X, Y, lat, lon, pred_id, fle_kg_km2, mcod_kg_km2, scod_kg_km2, date,
                bottom_depth) |>
  distinct(pred_id, .keep_all = TRUE) |>
  filter(predator_latin_name %in% "Gadus morhua") |>
  left_join(SE_wide3) |>
  rename(pred_weight = pred_weight_g,
         pred_length = pred_length_cm,
         Haul_ID = haul_id) |>
  mutate(Country = "SE",
         pred_weight_source = "observed")

```

## Merge old and new data

```{r merge dbs}

# rename ids and add a column to separate new from old database
new_db_data <- dd |> 
  rename(lat = ShootLat,
         lon = ShootLong,
         pred_weight = IndWgt) |>
  mutate(Haul_ID = as.character(tblHaulID),
         pred_ID = as.character(tblPredatorInformationID),
         data_source = "new_db")

old_db_data <- mm |> 
  mutate(data_source = "old_db",
         Day = as.double(Day),
         Month = as.double(Month))

new_SE_data <- ll |>
  rename(pred_ID = pred_id,
         Year = year,
         Depth = bottom_depth,
         ICESrectangle = ices_rect) |>
  dplyr::select(-c(X, Y)) |>
  mutate(data_source = "new_SE",
         Day = day(date),
         Month = month)

comcol_newold_pred <- intersect( intersect(colnames(new_db_data), colnames(old_db_data)),
                                 colnames(new_SE_data))

bind_rows( new_db_data |> dplyr::select(Year, pred_ID, data_source), 
           old_db_data |> dplyr::select(Year, pred_ID, data_source),
           new_SE_data |> dplyr::select(Year, pred_ID, data_source) ) |>
  group_by(Year, data_source) |>
  summarise(count = n()) |>
  ggplot() +
  geom_bar(aes(Year, count, fill = data_source), alpha = 1, stat="identity", position = "stack") +
  scale_x_continuous(n.breaks = 10) +
  ylab("# stomachs")

# Check if any of the two datasets have NA in the common columns. 
unique(is.na(old_db_data |> dplyr::select(all_of(comcol_newold_pred)))) # depth and time are all NAs and HaulNO we dont need. Day and month we may need however and these stomachs need to be removed in case month is a predictor
unique(is.na(new_db_data |> dplyr::select(all_of(comcol_newold_pred))))
unique(is.na(new_SE_data |> dplyr::select(all_of(comcol_newold_pred))))

str(old_db_data)
str(new_db_data)
# remove Time, Depth, HaulNo and Ship (ship makes w-o columns .x and .y)
newold <- full_join( old_db_data |> dplyr::select(-c(Time, Depth, HaulNo, Ship)), # join in predator data
                     new_db_data |> dplyr::select(-c(Time, Depth, HaulNo, Ship)),
                     by = comcol_newold_pred[!comcol_newold_pred %in% c("Time","Depth","HaulNo", "Ship")])  |>
  full_join(new_SE_data, by = comcol_newold_pred[!comcol_newold_pred %in% c("Time","Depth","HaulNo", "Ship")] )

```

Now calculate feeding ratios of cod. 

```{r}
# Calculate feeding ratios
newold2 <- newold |> 
  rowwise() |> 
  mutate(fr_sad = saduria / pred_weight,
         fr_spr = sprat / pred_weight,
         fr_her = herring / pred_weight,
         fr_other_or_NA = other_or_NA / pred_weight,
         fr_other_invert = other_invert / pred_weight,
         fr_other_fish = other_fish / pred_weight,
         fr_benth_fish = benth_fish / pred_weight,
         fr_tot = sum(other_or_NA, other_invert, benth_fish, other_fish, saduria, sprat, herring) / pred_weight)

newold2 |>  # log so we can see the data better
  dplyr::select(fr_sad, fr_spr, fr_her, fr_tot, fr_other_or_NA, fr_other_invert, fr_benth_fish) |> 
  pivot_longer(everything()) |> 
  filter(between(value, 0, 0.5 )) |>
  ggplot(aes(log(value))) +
  geom_histogram(bins = 100) +
  facet_wrap(~name, ncol = 2, scales = "free")
```

Check the proportion of stomachs without these prey

```{r}
newold2 |> 
  pivot_longer(c("herring", "saduria", "sprat","other_or_NA", "other_invert", "other_fish","benth_fish")) |> 
  group_by(name) |> 
  summarise(prop_empty = sum(value == 0)/n(), 
            prop_not_empty = sum(value != 0)/n())
```

Seems like fairly high proportions of cod without these key species in stomachs, but at least ~half have other prey in their stomachs

```{r}
#| out-width: 100%

# add day of the year
newold2 <- newold2 |> 
    mutate(day_of_year = ifelse( !is.na(sum(Year, Month, Day)),
                                  yday(as.Date(paste(Year, Month, Day, sep = "-"))),
                                  NA))

# Fix some columns
df <- newold2 |> 
  dplyr::select(!month) |> #remove errornous month
  rename(ices_rect = ICESrectangle,
         year = Year,
         month = Month,
         day = Day) |> 
  add_utm_columns(ll_names = c("lon", "lat"), utm_crs = 32633) |> 
  dplyr::select(pred_ID, herring, saduria, sprat, other_or_NA, other_invert, other_fish, benth_fish, fr_tot, fr_sad, fr_spr, fr_her, fr_other_or_NA, fr_other_invert, fr_benth_fish, fr_other_fish, year, month, day, day_of_year, pred_weight, pred_length, pred_weight_source, saduria, sprat, herring, lat, lon, ices_rect, X, Y, data_source, Country)#, depth)

# Add sample size per coordinate for plotting
df_plot <- df |>
  group_by(year, Y, X) |> 
  mutate(sample_size = n(),
         pos_id = paste(year, X, Y),
         decade = round(year/10) * 10) |> 
  ungroup() |> 
  distinct(pos_id, .keep_all = TRUE)

plot_map_fc +
  geom_point(data = df_plot, aes(X*1000, Y*1000, size = sample_size), alpha = 0.5) +
  #facet_wrap(~ year, ncol = 6) +
  facet_wrap(~ decade, ncol = 4) +
  #facet_wrap(Country ~ data_source , ncol = 6) +
  #theme_sleek(base_size = 6) + 
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  scale_size(range = c(.01, 2), name = "# stomachs") +
  geom_sf()

ggsave(paste0(home, "/figures/supp/year_diet_map.pdf"), width = 15, height = 15, units = "cm")
```

Add depth

```{r}
dep_raster <- terra::rast(paste0(home, "/data/Mean depth natural colour (with land).nc"))
class(dep_raster)
crs(dep_raster, proj = TRUE)

plot(dep_raster)

df$depth <- terra::extract(dep_raster, df |> dplyr::select(lon, lat))$elevation

ggplot(df, aes(lon, lat, color = depth*-1)) + 
  geom_point()

df$depth <- df$depth*-1

# TODO: these coordinates are waaay off
# This remoes 8349 rows
df <- df |> drop_na(depth) 

df |> 
  ggplot(aes(X*1000, Y*1000, color = depth)) + 
  geom_point() +
  NULL

hist(df$depth)

#remove points outside the baltic, i.e. the western most points in:
df |>
  ggplot(aes(lon, lat, color = depth)) + 
  geom_point() + 
  geom_vline(xintercept = 13) +
  coord_sf()

df <- df |>
  filter( lon > 13)

plot_map_fc +
  geom_point(data = df, aes(X*1000, Y*1000, color = depth), alpha = 0.5) +
  theme_sleek(base_size = 6) + 
  geom_sf()

```

```{r}
# Save data
write_csv(df, paste0(home, "/data/clean/stomachs.csv"))

```

```{r}
#| echo: false
#| eval: false
#renderthis::to_pdf(paste0(here::here(), "/R/prepare-data/01-test-prepare-stomach-data.html"))
```