---
title: "Compare smmother fits for prey weight by prey group"
author: "Viktor Thunell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 80%
editor: source
execute: 
  echo: true
  eval: true
  cache: true
---

## Load libraries

```{r libs}
#| message: false
#| warning: false
#| cache: false

# Load libraries, install if needed
pkgs <- c("tidyverse", "tidylog", "devtools", "sdmTMB", "sdmTMBextra", "terra", "mapplots",
          "viridis", "modelr", "beepr", "profmem") 

if(length(setdiff(pkgs,rownames(installed.packages()))) > 0){

    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  
  }

invisible(lapply(pkgs, library, character.only = T))

library(ggsidekick)
theme_set(theme_sleek())

# Set path
home <- here::here()
```


## Read stomach data

```{r}
#| message: false
#| warning: false
#| cache: false

#df <- read.csv(paste0(home, "/data/stomach/ran_rpwg_long.csv"))
df <- read.csv(paste0(home, "/QsdmTMB_dforum/Data_rpwg_long.csv"))

glimpse(df)

df2 <- df |>
  mutate(prey_group_f = as.factor(prey_group),
         val_cube = value^(1/3)) |>
  filter(between(year, 1985, 2015)) ##|> # reduce #obs by selecting years to make it possible to fit and predict models

df2 |>
  summarise(mean = mean(value), .by = c(year, prey_group)) |>
  ggplot(aes(year, mean, color = prey_group)) +
  geom_line() +
  facet_wrap(~prey_group) +
  expand_limits(x=c(1960,2020))

```

# Fit and check model

The number of observations are reduced to make them possible to work with (fitting and predicting). The data is not the original data as it is unpublished and has thus been altered. 

I´m excluding spatial and spatiotemporal random effects to prevent R from crashing (see my original spatial and spatiotemporal commented out). Contact me if you want ot try reproduce R crashing. 

```{r fit s}
#make mesh
mesh_long <- make_mesh(df2, c("X", "Y"), cutoff = 25)
plot(mesh_long)

# missing years
my <- min(df2$year):max(df2$year)
missing_years <- my[!my %in% unique(df2$year)]

# https://cran.r-project.org/web/packages/profmem/vignettes/profmem.html#:~:text=The%20profmem()%20function%20of,consumes%20more%20memory%20than%20expected.
capabilities("profmem")
#options(profmem.threshold = 300000)
#options(max.print=10000)

time <- Sys.time()
p <- profmem({
Mod_smooth_s <-
  sdmTMB(
  data = df2,
  mesh = mesh_long,
  formula = val_cube ~ 0 +
    prey_group_f +
    s(p_length_sc, year_sc, by = prey_group_f) + 
    s(d_sc, bs = "cc") +
    end_sc +
    ens_sc +
    ent_sc +
    eno_sc,
  extra_time = missing_years,
  spatial = "off",
  spatiotemporal = "off",
  # spatiotemporal = "iid",
  # spatial_varying = ~ 0 + prey_group_f,
  family = delta_lognormal(type = "poisson-link")
)
})

Sys.time() - time
```

The largest contribution to memory use is fitting smoothers.  

```{r check}
p |>
  filter(bytes > 5000000) |>
  arrange(-bytes)
total(p)/10^9 # gig

ifelse(exists("Mod_smooth_s"), sanity(Mod_smooth_s), "no fit")  # use ifelse exists when rendering and fitting may not be successful.

Mod_smooth_s

```


I can make predictions as the data is a subset (146 000) of the original 430 000 observations. 

```{r predict}
# predict on data
nd_Mod_test <- expand.grid("p_length_sc" = seq(min(df2$p_length_sc),
                                                  max(df2$p_length_sc),
                                                  length.out =30),
                            "prey_group_f" = levels(df2$prey_group_f),
                            "year_sc" = seq(min(df2$year_sc),
                                            max(df2$year_sc),
                                            length.out = length(unique(df2$year))))

nd_Mod_test <- nd_Mod_test |>
  mutate(depth_sc = 0,
         year = 1993,
         d_sc = 0,
         end_sc = 0,
         ens_sc = 0,
         ent_sc = 0,
         eno_sc = 0)

time <- Sys.time()
pred <- predict(Mod_smooth_s, newdata = nd_Mod_test, re_form = NA)
Sys.time() - time

format(object.size(pred), units = "auto")

```

# try fit the same model using a t2() smooth.
```{r fit t2}
time <- Sys.time()
  p2<-profmem({
  Mod_smooth_t2 <-
    sdmTMB(
    data = df2,
    mesh = mesh_long,
    formula = val_cube ~ 0 +
      prey_group_f +
      t2(p_length_sc, year_sc, by = prey_group_f, k=c(4,10)) +
      s(d_sc, bs = "cc") +
      end_sc +
      ens_sc +
      ent_sc +
      eno_sc,
    #time = "year",
    extra_time = missing_years,
    spatial = "off",
    spatiotemporal = "off",
    # spatiotemporal = "iid",
    # spatial_varying = ~ 0 + prey_group_f,
    family = delta_lognormal(type = "poisson-link")
  )
  })
  Sys.time() - time
```

Getting a model with the t2() smoother to fit is harder than with s(), also memory use is larger, fitting time longer. 

```{r check t2}
p2 |>
  filter(bytes > 10000000) |>
  arrange(-bytes)
total(p2)/10^9 # gig

ifelse(exists("Mod_smooth_t2"), sanity(Mod_smooth_t2), "no fit") # use ifelse exists when rendering and fitting may not be successful.
ifelse(exists("Mod_smooth_t2"), Mod_smooth_t2, "no fit")


```

